---
title: "Stringr"
author: "Anthony Arroyo"
date: "2022-11-18"
output:
  pdf_document: default
  html_document: default
---

```{r libs, include=FALSE}
library(magrittr)
library(stringr)
```

# Introduction 

Tidyverse has eight core packages according to its documentation.
These are the ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, and forcats packages.
Today we will investigate the [`stringr`](https://stringr.tidyverse.org/) package in more detail.
Stringr uses C libraries to manipulate string data and is built on stringi.

```{r}
data("HairEyeColor")
dimnames(HairEyeColor)
```

# Detect Matches

Stringr is able to detect patterns and count the number of occurrences.
It can check the beginning and end as well in one of its wrappers.

```{r detect}
colnames(HairEyeColor)

colnames(HairEyeColor) %>%
  stringr::str_starts(., "B")

colnames(HairEyeColor) %>%
  stringr::str_count(., "e")
```

# Mutate Strings

Stringr is able to change the value of a string based on a criteria.
For example, it can do substitution or convert all letters into uppercase.

```{r mutate}
rownames(HairEyeColor)

rownames(HairEyeColor) %>%
  stringr::str_replace(., "Brown", "Brunette")

rownames(HairEyeColor) %>%
  stringr::str_to_upper(.)
```

# Subset Strings

Stringr can produce subsets as well as substrings.
If we have a vector, we can match a pattern and only select those elements.
Also, we can take pieces of strings to reduce the length and abbreviate them.

```{r subset}
colnames(HairEyeColor)

colnames(HairEyeColor) %>%
  stringr::str_subset(., "B")

colnames(HairEyeColor) %>%
  stringr::str_sub(., 1, 2)
```

# Order and Sort

Finally, and perhaps most useful of all, we can use stringr to sort strings.
It will do any hexcode wrangling under the hood so that you don't have to.

```{r sort}
rownames(HairEyeColor)

rownames(HairEyeColor) %>%
  stringr::str_order(.)

rownames(HairEyeColor) %>%
  stringr::str_sort(.)

rownames(HairEyeColor) %>%
  stringr::str_sort(., decreasing = TRUE)
```

