---
title: "TidyVerse_Extend_Assignment"
author: "Gabriel Santos"
date: '2022-11-11'
output: html_document
---
Source: https://www.kaggle.com/ronitf/heart-disease-uci/data#heart.csv


```{r}
library(tidyverse)
```

```{r}
heartdata <- read.csv('https://raw.githubusercontent.com/GabrielSantos33/DATA-607_Tidyverse_Extend/main/HeartDisease.csv')
view(heartdata)
head(heartdata)
```


```{r}
filteredheart <- heartdata%>%
  filter(Cholesterol > 200)
head(filteredheart)
```


```{r}
selectheart <- heartdata%>%
  select(c('Age','MaxHR','HeartDisease'))
head(selectheart)
```


```{r}
bysex <- heartdata%>%
  group_by(Sex)%>%
  summarise(count = n())

bysex
```

### *Part 2: Tidyverse Extend*

dplyr:: select

Q: How do i select from one column to another in a table

A: By using the dplyr select function

```{r}
heartdata2 <- heartdata%>%
  select(Age:Cholesterol,HeartDisease)
head(heartdata2)
```

dplyr::rename

Q: How do i rename a variable in a table?

A: I use dplyr rename function

```{r}
heartdata2 <- heartdata2 %>% rename("Gender" = Sex)
head(heartdata2)
```

dplyr::case_when()

Q: How do i vectorise multiple if else statement

A: By Using the case_when() function

```{r}
heartdata3 <- heartdata2 %>% 
  mutate(Heart_condition = case_when(Cholesterol >= 300 ~ 'Severe'
                                     ,TRUE ~ 'Not_severe' )
         )
head(heartdata3)
```

ggplot2::Barplot

Q: I want to plot a Barchart of the count of Male and Female in the dataset

A: Use the ggplot function and the geom_bar layer
```{r}
ggplot(data = heartdata2, aes(x = Gender)) + geom_bar(fill = "blue") + 
  labs(title = "Bar chart for count of Sex") + theme_bw()
```
coord_flip()

Q: How do I flip co-ordinates?

A: Add the coord_flip layer

```{r}
ggplot(data = heartdata2, aes(x = Gender)) + geom_bar(fill = "blue") + 
  labs(title = "Bar chart for count of Sex") + theme_bw() + coord_flip()
```

ggplot: Histogram
```{r}
# Histogram of Cholesterol level
ggplot(heartdata2, aes(x = Cholesterol)) + geom_histogram(binwidth = 30, fill = "darkblue") + 
  labs(title = "Distribution of Cholesterol level", subtitle = "Cholesterol level") + ylab(NULL)
```


