---
title: ' DATA 607 Tidyverse Assignment'
author: "Gabriel Santos"
date: "2022-10-29"
output:
  html_document: default
  pdf_document: default
---

***Instruction***

Create an Example TidyVerse packages, dataset from Kaggle. 
Created sample  program “vignette” demonstrating how to use one or more of the abilities of the selected TidyVerse package in the dataset.


```{r}
library(tidyverse)
library(ggplot2)
library (dplyr)
```


***Columns***

age:age in years.
sex:(1 = male; 0 = female)
cpchest: pain type. Type of chest-pain experienced by the individual:
1 = typical angina
2 = atypical angina
3 = non-angina pain
4 = asymptomatic angina
trestbps: Resting blood pressure (in mm Hg on admission to the hospital).
chol:Serum cholestoral in mg/dl.
fbs:(fasting blood sugar > 120 mg/dl) (1 = true; 0 = false).
restecg: Resting electrocardiographic results:
0 = normal
1 = ST-T wave abnormality
2 = left ventricle hyperthrophy
thalach: Maximum heart rate achieved.
exang: Exercise induced angina (1 = yes; 0 = no).
oldpeak: ST depression induced by exercise relative to rest.
slope: the slope of the peak exercise ST segment.
ca: number of major vessels (0-3) colored by flourosopy.
thal: 3 = normal; 6 = fixed defect; 7 = reversable defect.
target: 1 or 0.


Source: https://www.kaggle.com/ronitf/heart-disease-uci/data#heart.csv

```{r}
Heart <- read.csv("https://raw.githubusercontent.com/GabrielSantos33/DATA-607-Tidyverse/main/HeartDisease.csv")
head(Heart)
```


```{r}
names (Heart) <- c("Age","Gender", "Chest_Pain_Type","Resting_Blood_Pressure","Serum_Cholesterol","Fasting_Blood_Sugar","Resting_ECG","Max_Heart_Rate_Achieved","Exercise_Induced_Angina","ST_Depression_Exercise","Peak_Exercise_ST_Segment","Num_Major_Vessels_Flouro", "Thalassemia","Diagnosis_Heart_Disease")
head(Heart)
```

Display the numbers of rows and columns in the dataset.


```{r}
dim(Heart)
```
The data contain 303 observation and 12 vaiables.


```{r}
str(Heart)
```

```{r}
sum(is.na(Heart))
```
```{r}
tibble <- as_tibble(Heart)
```


```{r}
Heart$Age <- as.factor(Heart$Age)
Heart$Gender <- as.factor (Heart$Gender)
Heart$Diagnosis_Heart_Disease <- as.factor (Heart$Diagnosis_Heart_Disease)
```


```{r}
summary(Heart)
```

The summary function explains alot of things about the data. In this data, we can have some knowledge about our data features and observation. For example, minimum, maximum, mean, median, quarters. 


```{r}
Heart %>%
  summarise (Mean = mean(Resting_Blood_Pressure ), Max = max(Resting_Blood_Pressure), Mean = mean(Resting_Blood_Pressure ), Variance= var(Resting_Blood_Pressure ), SD= sd(Resting_Blood_Pressure))
```

```{r}
ggplot(Heart,aes(x= Num_Major_Vessels_Flouro,fill= Diagnosis_Heart_Disease)) +
  geom_bar()+ labs(y ="count", title = "Heart disease diagnosis", subtitle = "Number of major vessels")
```


This graph shows that major vessel 0 causes heart diseases the most.


```{r}
ggplot(Heart,aes(x= Max_Heart_Rate_Achieved,fill= Diagnosis_Heart_Disease)) +
  geom_bar()+ labs(y ="count",title = "Heart disease diagnosis", subtitle = "Max Heart Rate Achieved")
```

Heart rate is also important in diagnosis of heart disease



```{r}
ggplot(Heart,aes(x= Gender,fill= Diagnosis_Heart_Disease)) +
  geom_bar()+ labs(y ="Age", title = "Heart disease diagnosis ", subtitle = "Rate distribution by Gender and Age")
```


This graph show that males are diagnosed with heart disease more than females. Gender:(1 = male; 0 = female)



```{r}
ggplot(Heart,aes(x= Chest_Pain_Type,fill= Diagnosis_Heart_Disease)) +
  theme_bw() +  geom_bar() +
  facet_wrap(~Gender) +
  labs(y ="count", title = "Heart Disease distribution", subtitle = "Gender based on Chest_Pain_Type")
```

This graph shows that people (males and females) experience chest pain (2), which is (non-angina pain) have higher chance of having a heart disease.This graph also shows that Males have higher exposure to have heart diseases if they suffer from (typical angina, atypical angina, or asymptomatic angina)



```{r}
ggplot(Heart,aes(x= Resting_Blood_Pressure,fill= Diagnosis_Heart_Disease)) +
  geom_bar()+
   labs(y ="count", title = "Heart Disease diagnostic Rates", subtitle = "Resting_Blood_Pressure")
```


```{r}
ggplot(Heart,aes(x= Serum_Cholesterol,fill= Diagnosis_Heart_Disease)) +
  geom_bar()+ labs(y ="Frequency", title = "Heart Disease Diagnostic Rates", subtitle = "Serum Cholesterol")
```

```{r}
ggplot(Heart, aes(ST_Depression_Exercise, Resting_Blood_Pressure, colour = Thalassemia)) + 
    geom_point() + ggtitle("Heart", subtitle = "Analysis")
```

***Conclusion***

*Heart disease is a significant health concern, there are factors that people should be aware of. In addition to age, Chest Pain, Cholesterol, and Blood Pressure are essential factors to bear in mind.*
